### 2018-01-22 slides/rpctrace.tex
add slides describing rpctrace "attach" operation

\documentclass{beamer}
\usetheme{Madrid}

\setbeamertemplate{footline}{}
\beamertemplatenavigationsymbolsempty

\usepackage{comment}
\usepackage{hurd}

\usetikzlibrary{positioning, calc}

\begin{document}

\tikzstyle{send} = [rectangle, fill=green, rounded corners] 
\tikzstyle{recv} = [rectangle, fill=red!50, rounded corners]
\tikzstyle{empty} = [minimum size=0, inner sep=0]
\tikzstyle{queue} = [draw, rectangle, minimum height=.5cm, minimum width=1cm, pattern=vertical lines]

\begin{frame}
\begin{tikzpicture}

\node (center point) {};

\task [left = of center point] {rpctrace}{}{lr}{}
\task [right = of center point] {PUT}{}{l}{}

\queue [above = of rpctrace] {queue}{}

\node (below rpctrace) [below = of rpctrace] {};
\node (below PUT) [below = of PUT] {};

% invisible at first, but present to keep the canvas sizing consistent across all slides
\only<1-4>{
  \queue [white] {queue2}{at ($(below rpctrace)!.5!(below PUT)$)}
}
\only<5->{
  \queue {queue2}{at ($(below rpctrace)!.5!(below PUT)$)}
}

% invisible receive right to match one from the next set of slides, so the canvas sizing is identical
\draw [white, thick] (queue.west) -- ++(-3,1) node [recv, white, anchor=south] {R};

\only<1,8->{\node (R1) at (PUT Port 0) [recv, anchor=west] {R};}
\only<2-7>{\node (R1) at (PUT Port 0) [recv, anchor=west, minimum size=0, inner sep=0] {};}
\node (r1) at (R1) {};

\only<1-3,7->{
  \node (S1) at (R1.east) [send, anchor=west] {S};
  \node (S2) at (S1.east) [send, anchor=west] {S};
}

\only<1-3>{
  \draw [violet, thick] (PUT Port 0) -- +(-1,0) |- (queue.east);
}

\draw [violet, thick] (queue.north) -- ++(-1,1) node [send, anchor=south] {S};
\draw [violet, thick] (queue.north) -- ++(1,1) node [send, anchor=south] {S};

\only<2->{
  \node (R2) at (rpctrace Left Port 0) [recv, anchor=west] {R};
  \draw [violet, thick] (queue.west) -- ++(-2,0) |- (R2);
}

\only<3-9>{
  \node (S3) at (R2.east) [send, anchor=west] {S};
}
\only<10->{
  \node (S3) at (R2.east) [empty] {};
}

\only<5-7>{
  \node (R3) at (rpctrace Right Port 0) [recv, anchor=east] {R};
}
\only<8->{
  \node (R3) at (rpctrace Right Port 0) [minimum size=0, inner sep=0] {};
}

\only<5->{
  \draw [violet, thick] (rpctrace Right Port 0) -| ($(rpctrace Right Port 0)!.5!(queue2.west)$) |- (queue2.west);
}

\only<6->{
  \node (S4) at (R3.west) [send, anchor=east] {S};
}

\only<7->{
  \draw [violet, thick] (queue2.east) -| ($(queue2.east)!.5!(PUT Port 0)$) |- (PUT Port 0);
}

\only<9->{
  % S3's location moves between slides, so starting this line at S4 keeps the dashes from morphing between slides
  \draw [violet, dashed, thick] (S4) -- (S3);
}

\end{tikzpicture}
\end{frame}

\begin{comment}
    (defun increment-number-at-point ()
      (interactive)
      (skip-chars-backward ``0-9'')
      (or (looking-at ``[0-9]+'')
          (error ``No number at point''))
      (replace-match (number-to-string (1+ (string-to-number (match-string 0))))))

(when (string-match ``[ \t]*$'' test-str)
    (message (concat ``[`` (replace-match " nil nil test-str) ``]'')))

     (while (re-search-forward ``foo[ \t]+bar'' nil t)
       (replace-match ``foobar''))

(defun ff ()
  "increment numbers in string"
  (let (mystr start)
    (setq str (match-string 0)
    (setq start 0)
    (while (string-match "[0-9]+" mystr start)
      (replace-match (number-to-string (1+ (string-to-number (match-string 0)))) nil nil (match-string 0))
  (replace-regexp-in-string ``_'' `` `` (match-string 1))
)

\end{comment}

\begin{frame}
\begin{tikzpicture}

\node (center point) {};

\task [left = of center point] {rpctrace}{}{lr}{}
\task [right = of center point] {PUT}{}{l}{}

\queue [above = of rpctrace] {queue}{}

\node (below rpctrace) [below = of rpctrace] {};
\node (below PUT) [below = of PUT] {};

% invisible at first, but present to keep the canvas sizing consistent across all slides
\only<1-3>{
  \queue [white] {queue2}{at ($(below rpctrace)!.5!(below PUT)$)}
}
\only<4->{
  \queue {queue2}{at ($(below rpctrace)!.5!(below PUT)$)}
}

\only<1-2,5->{
  \node (S1) at (PUT Port 0) [send, anchor=west] {S};
  \node (S2) at (S1.east) [send, anchor=west] {S};
}

\only<1-2>{
  \draw [violet, thick] (PUT Port 0) -- +(-1,0) |- (queue.east);
}

\draw [violet, thick] (queue.north) -- ++(-1,1) node [send, anchor=south] {S};
\draw [violet, thick] (queue.north) -- ++(1,1) node [send, anchor=south] {S};
\draw [violet, thick] (queue.west) -- ++(-3,1) node [recv, anchor=south] {R};

\only<2->{
  \node (S3) at (rpctrace Left Port 0) [send, anchor=west] {S};
  \draw [violet, thick] (queue.west) -- ++(-2,0) |- (S3);
}

\only<4->{
  \node (R3) at (rpctrace Right Port 0) [recv, anchor=east] {R};
}

\only<4->{
  \draw [violet, thick] (rpctrace Right Port 0) -| ($(rpctrace Right Port 0)!.5!(queue2.west)$) |- (queue2.west);
}

\only<5->{
  \draw [violet, thick] (queue2.east) -| ($(queue2.east)!.5!(PUT Port 0)$) |- (PUT Port 0);
}

\only<6->{
  \draw [violet, dashed, thick] (S3) -- (R3);
}

\end{tikzpicture}
\end{frame}

\end{document}